---
title: "Class II Assignment"
author: "Akhil Havaldar"
date: "9/1/2021"
output: html_document
---

```{r}
library(tidyverse)
```
# Data Set 1
```{r}
nfl_dat <- read_csv("/Users/a8hav/Downloads/nfl2008_fga.csv")
```
## Question 1: Are there differences in the field goal distance by down for attempts in 2008?

## Pseudocode 1
    - Group by down
    - calculate average distance per down
    - arrange in order of descending avg distance

## Single Line Version 1   
```{r}
# Group by down
as.factor(nfl_dat$down)
x1 <- group_by(nfl_dat, down)

# Calculate averages
x2 <- summarize(x1, count=n(), mean(distance))

# Arrange by averages
x3 <- arrange(x2, desc(`mean(distance)`))
```
## Piped Version 1
```{r}
as.factor(nfl_dat$down)
x1 <- group_by(nfl_dat, down) %>%
  summarize(count=n(), mean(distance)) %>%
  arrange(desc(`mean(distance)`))
```
## Function 1
```{r}
avg_dist <- function(x){
  a1 <- group_by(x, down) %>%
    summarize(count=n(), mean(distance)) %>%
    arrange(desc(`mean(distance)`))
  print(a1)
}

avg_dist(nfl_dat)
```

## Question 2: Which kickers were the most successful in kicking field goals over 40 yards?

## Pseudocode 2
      - Group by kicker
      - filter by FG attempts of over 40 yards
      - count the number of attempts and makes
      - calculate FG percentage for each kicker
      
## Single line 2
```{r}
x1 <- group_by(nfl_dat, name)
x2 <- filter(x1, distance>40)
x3 <- summarize(x2, attempts = n(), makes = sum(GOOD))
x4 <- mutate(x3, fgpct = makes / attempts)
x5 <- arrange(x4, desc(`fgpct`))

```
## Piped Version 2
```{r}
x1 <- group_by(nfl_dat, name) %>%
  filter(distance>40) %>%
  summarize(attempts = n(), makes = sum(GOOD)) %>%
  mutate(fgpct = makes/attempts)%>%
  arrange(desc(`fgpct`))
```
## Function 2
```{r}
fgpct <- function(x){
  a1 <- group_by(x, name) %>%
  filter(distance>40) %>%
  summarize(attempts = n(), makes = sum(GOOD)) %>%
  mutate(fgpct = makes/attempts)%>%
  arrange(desc(`fgpct`))
  
  print(a1)
}

fgpct(nfl_dat)
```


## Data Set 2
```{r}
library(readr)
beaches <- read_csv("~/DS 3001/DS-3001/data/beaches.csv")
```
## Question 1: Which months have higher levels of enterococci on average for temps greater than 25?

## Pseudocode 1:
      - Group by month
      - filter by temps greater than 25
      - calculate average enterococci 
      - arrange by descending values
      
## Single line code 1
```{r}
x1 <- group_by(beaches, month_name)
x2 <- filter(x1, temperature>25)
x3 <- summarize(x2,mean = mean(enterococci))
x4 <- arrange(x3, desc(mean))
```

## Piped Version 1:
```{r}
x1<-group_by(beaches, month_name)%>%
  filter(temperature>25)%>%
  summarize(mean=mean(enterococci))%>%
  arrange(desc(mean))
```
## Function 1
```{r}
mean_entero <- function(x){
  a1 <- group_by(x, month_name)%>%
  filter(temperature>25)%>%
  summarize(mean=mean(enterococci))%>%
  arrange(desc(mean))
  
  print(a1)
}

mean_entero(beaches)
```

## Question 2: What year was rainfall the most prominent in Spring?

## Pseudocode 2:
    - group by year
    - filter by spring
    - calculate average rainfall
    - sort by descending values

## Single Line 2:
```{r}
x1 <- group_by(beaches, year)
x2 <- filter(x1, season_name == "Spring")
x3 <- summarize(x2, mean = mean(rainfall, na.rm = T))
x4 <- arrange(x3, desc(mean))
```

## Piped Version 2:
```{r}
x1 <- group_by(beaches, year)%>%
  filter(season_name == "Spring")%>%
  summarize(mean = mean(rainfall, na.rm = T))%>%
  arrange(desc(mean))
```

## Function 2:
```{r}
rain_mean <- function(x){
  a1 <- group_by(x, year)%>%
    filter(season_name == "Spring")%>%
    summarize(mean = mean(rainfall, na.rm = T))%>%
    arrange(desc(mean))
  print(a1)
}

rain_mean(beaches)
```

## Data Set 3
```{r}
library(readr)
wine <- read_csv("~/DS 3001/DS-3001/data/winequality-red-ddl.csv")
names(wine)[1]<-"fixed_acidity"
names(wine)[c(2,3)]<-c("vol_acid", "citric_acid")
```
## Question 1: Does the fixed acidity of wine impact how highly the wine is rated?

## Pseudocode 1:
      - Group by wine quality (text)
      - calculate the average fixed acidity per each wine quality
      - sort by descending values to determine rank

## Single line 1:
```{r}
x1 <- group_by(wine, text_rank)
x2 <- summarize(x1, mean = mean(fixed_acidity, na.rm=T))
x3 <- arrange(x2, desc(mean))
```
## Piped Version 1: 
```{r}
x1<-group_by(wine, text_rank)%>%
  summarize(mean = mean(fixed_acidity, na.rm=T))%>%
  arrange(desc(mean))
  
```
## Function 1:
```{r}
fixed_acid_avg <- function(x){
  a1 <- group_by(x, text_rank)%>%
    summarize(mean = mean(fixed_acidity,na.rm=T))%>%
    arrange(desc(mean))
  print(a1)
}
fixed_acid_avg(wine)
```

## Question 2: For high quality wines, is citric acid or volatility acid higher on average?

## Pseudocode 2:
      - Select only the "excellent" wines
      - average the volatility acid
      - average the citric acid
      - merge the datasets
      
## Single Line 2:
```{r}
x1 <- filter(wine, text_rank == "excellent")
x2 <- summarize(x1, mean_vol = mean(vol_acid))
x3 <- summarize(x1, mean_cit = mean(citric_acid))
x4 <- merge(x2, x3)

```
## Piped Version 2: 
```{r}
x1 <- filter(wine, text_rank == "excellent")%>%
  summarize(mean_vol = mean(vol_acid))
x2<- filter(wine, text_rank == "excellent")%>%
  summarize(mean_cit = mean(citric_acid))
x3<-merge(x1, x2)
```

## Function 2:
```{r}
func <- function(x){
  a1 <- filter(x, text_rank == "excellent")%>%
    summarize(mean_vol = mean(vol_acid))
  a2 <- filter(x, text_rank == "excellent")%>%
  summarize(mean_cit = mean(citric_acid))
  a3 <- merge(a1, a2)
  print(a3)
}
func(wine)
```

